---
title: "HW5"
author: "Amane Chibana and Harry Wang"
date: "2024-03-03"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(BSDA)
library(EnvStats)
```


### Question 1

##### i)

$\sigma_{1}^{2}\ =\ \sigma_{2}^{2}$ The two populations have equal variances

$\sigma_{1}^{2}\ne\sigma_{2}^{2}$ The two populations do not have equal variances


##### ii)

```{r}
organic = c(4.83,5.50,6.33,6.33,5.67,6.17,5.50,5.00,5.00,6.33,6.17,4.83,4.67,5.50,5.83,4.67,6.17,5.33,6.17,5.67)
control = c(5.33,5.67,5.33,5.17,5.67,5.67,3.83,5.33,4.67,4.83,5.33,3.83,5.33,5.33,4.33,5.50,5.17,6.17,4.83,4.33)
comfort = c(5.83,5.00,6.00,5.00,4.17,4.67,5.17,5.50,4.67,4.50,5.17,5.17,4.50,4.67,4.83,5.83,5.17,4.33,4.67,4.33,4.67,3.67)
test_stat <- (sd(control)^2)/(sd(comfort)^2)
n1 = length(control)
n2 = length(comfort)
```

```{r, echo=FALSE}
cat("Testing Statistic:\n")
cat(test_stat, "\n")

cat("The null distribution is a F distribution of with degree of freedom ", n1-1, " (# of control - 1) and ", n2-1, " (# of comfort - 1)")
```

##### iii)

```{r}
var.test(control ,comfort , ratio =1,alternative =c("two.sided"))
```

```{r,echo=FALSE}
cat("We fail to reject the null hypothesis because the p-value is greater than the significance value of 0.05.")
```

##### iv)

```{r,echo=FALSE}
cat("This means that we lack the evidence to conclude that the null hypothesis is invalid. It does not confirm that the null hypothesis is true but there is a lack of evidence to say that the populations have different variances.")
```

### Question 2


##### i)

$\sigma_{1}^{2}\ =\ \sigma_{2}^{2}$ The two populations have equal population means

$\sigma_{1}^{2}\ne\sigma_{2}^{2}$ The two populations do not have equal population means


##### ii)

```{r}
organic = c(4.83,5.50,6.33,6.33,5.67,6.17,5.50,5.00,5.00,6.33,6.17,4.83,4.67,5.50,5.83,4.67,6.17,5.33,6.17,5.67)
control = c(5.33,5.67,5.33,5.17,5.67,5.67,3.83,5.33,4.67,4.83,5.33,3.83,5.33,5.33,4.33,5.50,5.17,6.17,4.83,4.33)
comfort = c(5.83,5.00,6.00,5.00,4.17,4.67,5.17,5.50,4.67,4.50,5.17,5.17,4.50,4.67,4.83,5.83,5.17,4.33,4.67,4.33,4.67,3.67)

ss1=var(control); ss2=var(comfort)

n1=length(control); n2=length(comfort)

t=( mean(control)-mean(comfort))/sqrt(ss1/n1+ss2/n2)
```

```{r, echo=FALSE}
cat("Testing Statistic:",t)

cat("The null distribution is a T distribution with degree of freedom ", n1+n2-2, " (# of control + # of comfort - 2)")
```

##### iii)

```{r}
t.test(control,comfort,alternative =c("two.sided"), paired =FALSE ,var.equal = TRUE)
```

```{r, echo=FALSE}
cat("Since the p-value (0.2959) is greater than the typical significance level of 0.05, we fail to reject the null hypothesis.")
```

##### iv)

```{r, echo=FALSE}
cat("This means there is not enough statistical evidence to conclude that there is a significant difference in the means of the two populations.")
```     


### Question 3

#### i)

$p_1 = p_2$ The two populations have equal population proportions.

$p_1\ne p_2$ The two populations do not have equal population proportions.

#### ii)

```{r}
organic = c(4.83,5.50,6.33,6.33,5.67,6.17,5.50,5.00,5.00,6.33,6.17,4.83,4.67,5.50,5.83,4.67,6.17,5.33,6.17,5.67)
control = c(5.33,5.67,5.33,5.17,5.67,5.67,3.83,5.33,4.67,4.83,5.33,3.83,5.33,5.33,4.33,5.50,5.17,6.17,4.83,4.33)
comfort = c(5.83,5.00,6.00,5.00,4.17,4.67,5.17,5.50,4.67,4.50,5.17,5.17,4.50,4.67,4.83,5.83,5.17,4.33,4.67,4.33,4.67,3.67)

controlbinary = ifelse(control>4.75,1,0)
comfortbinary = ifelse(comfort>4.75,1,0)

mean1 = mean(controlbinary)
mean2 = mean(comfortbinary)
ptotal = (15+11)/(20+22)
z = (mean1-mean2)/(sqrt(ptotal*(1 - ptotal)*((1/20)+(1/22))))
```

```{r, echo=FALSE}
cat("Testing Statistic:",z)

cat("The null distribution is a Z (Normal) distribution.")
```

#### iii)
```{r}
successes = c(15,11)
total= c(20,22)
prop.test(successes,total, alternative =c("two.sided"),conf.level=0.95, correct=FALSE)
```

```{r, echo=FALSE}
cat("Since the p-value (0.0957) is greater than the typical significance level of 0.05, we fail to reject the null hypothesis.")
```

#### iv)

```{r, echo=FALSE}
cat(" Since the P-value is greater than the common alpha level of 0.05, we do not reject the null hypothesis. This suggests that there is not enough statistical evidence to say there is a significant difference between the proportions of observations greater than 4.75 in the Control and Comfort groups.")
```     