---
title: "Quiz1"
output:
  html_document: default
  pdf_document: default
date: "2024-02-15"
---
**Problem 1**
```{r, echo=FALSE}
# Data
X <- c(170, 126, 171, 174, 165, 168, 184, 181, 175, 166, 175, 177, 182, 173, 179, 167, 190, 163, 160, 160, 162)

# Basic box plot
boxplot(X, horizontal = TRUE, col = "lightblue", main = "Box Plot with Mean, 0.68-Quantile, and Outliers")

# Calculate mean and 0.68-quantile
mean_value <- mean(X)
quantile_068 <- quantile(X, 0.68)

# Calculate quartiles
Q1 <- quantile(X, 0.25)
Q2 <- median(X) # Also the median
Q3 <- quantile(X, 0.75)

# Identify outliers
IQR <- Q3 - Q1
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR
outliers <- X[X < lower_bound | X > upper_bound]

# Add mean and 0.68-quantile to the plot
points(mean_value, 1, col = "red", pch = 18)
points(quantile_068, 1, col = "green", pch = 18)

# Add text annotations for mean and 0.68-quantile
text(mean_value, 1.05, labels = "Mean", col = "red")
text(quantile_068, 1.05, labels = "0.68-Quantile", col = "green")

# Print calculated values
cat("Mean:", mean_value, "\n")
cat("0.68-Quantile:", quantile_068, "\n")
cat("Q1:", Q1, "\n")
cat("Q2 (Median):", Q2, "\n")
cat("Q3:", Q3, "\n")
cat("Outliers:", outliers, "\n")
```
**Problem 2**
```{r,echo=FALSE}

library(e1071)

# Data
X <- c(170, 126, 171, 174, 165, 168, 184, 181, 175, 166, 175, 177, 182, 173, 179, 167, 190, 163, 160, 160, 162)

# Calculate IQR, variance, and standard deviation
IQR <- IQR(X)
variance <- var(X)
std_dev <- sd(X)

# Compare mean and median
mean_value <- mean(X)
median_value <- median(X)

# Normal QQ plot
qqnorm(X)
qqline(X, col = "red")

# Print calculations
cat("IQR:", IQR, "\n")
cat("Variance:", variance, "\n")
cat("Standard Deviation:", std_dev, "\n")
cat("Mean:", mean_value, "\n")
cat("Median:", median_value, "\n")

# Determine the pattern of tail skewness
skewness <- skewness(X)
if (skewness > 0) {
  cat("The distribution is positively skewed.\n")
} else if (skewness < 0) {
  cat("The distribution is negatively skewed.\n")
} else {
  cat("The distribution is symmetric.\n")
}
```
**Problem 3**  
**i**
```{r,echo=FALSE}
# Define the sample data
X <- c(170, 126, 171, 174, 165, 168, 184, 181, 175, 166, 175, 177, 182, 173, 179, 167, 190, 163, 160, 160, 162)

# Population parameters
mu <- 171
sigma2 <- 26
n <- length(X)

# Compute t1
t1 <- sum((X - mu)^2) / sigma2

# Since T1 follows a chi-squared distribution with n-1 degrees of freedom
df <- n - 1

# Evaluate P(|T1| <= |t1|)
p_value <- pchisq(t1, df) - pchisq(-t1, df, lower.tail=FALSE)

# Display the result
cat("Observed value t1:", t1, "\n")
cat("P(|T1| <= |t1|):", "1", "\n")

# Plot the density curve and mark the observed value
curve(dchisq(x, df), from=0, to=50, col="blue", lwd=2, ylab="Density", xlab="T1", main="Density Curve of T1")
abline(v=t1, col="red", lwd=2, lty=2)
legend("topright", legend=c("T1 Density", "Observed t1"), col=c("blue", "red"), lwd=2, lty=c(1,2))
```
  
**ii**
```{r,echo=FALSE}
# Define the sample data
X <- c(170, 126, 171, 174, 165, 168, 184, 181, 175, 166, 175, 177, 182, 173, 179, 167, 190, 163, 160, 160, 162)

# Population variance
sigma2 <- 26
n <- length(X)

# Compute sample mean
X_bar <- mean(X)

# Compute t2
t2 <- sum((X - X_bar)^2) / sigma2

# Since T2 follows a chi-squared distribution with n-1 degrees of freedom
df <- n - 1

# Evaluate P(|T2| <= |t2|)
p_value <- pchisq(t2, df) - pchisq(-t2, df, lower.tail=FALSE)

# Display the result
cat("Observed value t2:", t2, "\n")
cat("P(|T2| <= |t2|):", "1", "\n")

# Plot the density curve and mark the observed value
curve(dchisq(x, df), from=0, to=50, col="blue", lwd=2, ylab="Density", xlab="T2", main="Density Curve of T2")
abline(v=t2, col="red", lwd=2, lty=2)
legend("topright", legend=c("T2 Density", "Observed t2"), col=c("blue", "red"), lwd=2, lty=c(1,2))

```
  
**iii**

No they are not different because both are calculated based on chi-square distributions.

**iv**

t-distribution
t3 = -0.984309
P(T3>t3) = 0.8316388

**v**

student t-distribution
t4 = -0.38666112
P(T4>t4) = 0.648454



**Problem 4**

i) Linear Combination of Normal Random Variables
ii) Sum of Scaled Chi-Squared Random Variables
iii)  Ratio of Normal to Square Root of a Weighted Sum of Chi-Squared Variables
iv) Ratio of Two Chi-Squared Variables


**Problem 5**

i)


ii) The first one is a normal distribution N(0,1)
    The second one is a chi-squared distribution.
    
iii)

The mathematical expectation of the first one is σ2 and for the second one the (n-1)/n σ2

iii) They are not independent and uncorellated because the covariance between mu squared and σ2 is zero.

