---
title: "Midterm"
author: "Harry Wang"
date: "`r Sys.Date()`"
output: html_document
---

I pledge my honor that I have abided by the Stevens Honor System

#### Problem 1:

```{r}
library(readxl)
GpaGender = read_excel("./GpaGender.xls")
n = length(GpaGender$SelfConcept)
n
significance = 0.05

#males
X = GpaGender[which(GpaGender$Gender == 2), "SelfConcept"]
X
male_summary = summary(X)
male_summary

#females
Y = GpaGender[which(GpaGender$Gender == 1), "SelfConcept"]
Y
female_summary = summary(Y)
female_summary

```

#### Problem 2:

```{r}
together = c(X,Y)
together
boxplot(together,horizontal = TRUE, names = c("males","females"))
```

#### Problem 3:

```{r}
t.test(X,conf.level = 0.95)
```

The 95% confidence interval for the true male population mean is 54.31 to 61.52.

```{r}
t.test(Y,conf.level = 0.95)
```

The 95% confidence interval for the true female population mean is 50.86 to 60.17.

#### Problem 4:

```{r}

cat("The null hypothesis is H0 : σ2X = σ2Y")
nx = length(X$SelfConcept)
nx
ny = length(Y$SelfConcept)
ny
var(X$SelfConcept)
var(Y$SelfConcept)
```

The sample variance of sample X (male) is less than that of sample Y (female). So, I set my alternative hypothesis to be Ha : σ2X < σ2Y

```{r}
var.test(X$SelfConcept, Y$SelfConcept, ratio = 1, alternative = "less")

```
The testing statistic F is observed as f = 0.93

The p-value 0.41 is not less than the significance level 0.05, so I fail to reject the null hypothesis, meaning that the
male population Self Concept variance is equal to the female population Self Concept variance.

#### Problem 5:

The null hypothesis is H0 : μX = μY
The alternative hypothesis is Ha : μX != μY

```{r}
t.test(X$SelfConcept,Y$SelfConcept, alternative = "two.sided", paired = FALSE, var.equal = TRUE)
```

The p-value 0.41 is not less than the significance level 0.05, so I fail to reject the null hypothesis, meaning
that the male population Self Concept mean is equal to the female population Self Concept mean.

#### Problem 6:

The null hypothesis is H0 : p = 0.45
The alternative hypothesis is Ha : p > 0.45

```{r}
bin = ifelse(GpaGender$SelfConcept > 50, 1 ,0)
bin

prop.test(sum(bin), length(bin), p = 0.45, alternative = "greater", correct = FALSE)
```
The p-value 1.876e-09 is less than the significance level 0.05, so I reject the null hypothesis, meaning that
the population proportion is greater than 0.45.

#### Problem 7:

The null hypothesis is H0 : p1 = p2
The alternative hypothesis is Ha : p1 !- p2       

```{r}
malebin = ifelse(X > 50, 1, 0)
femalebin = ifelse(Y > 50,1,0)

prop.test(c(sum(malebin),sum(femalebin)), c(length(malebin),length(femalebin)), alternative = "two.sided", correct = FALSE)
```


The p-value 0.4858 is greater than the significance level 0.05, so I fail to reject the null hypothesis, meaning that
the two groups have the same population proportion of Self Concept through a two tailed test.